---
title: "R Notebook"
output: html_notebook
---


```{r message=FALSE, warning=FALSE}
library(tidyverse)
```


Payoff function
```{r}

R0 <- 2

#pi is the probability (0~1) of being infected in the steady state with vaccination fraction of p.
pi <- function(p = 0.5, R0 = 2) {
  val = 1 - (1/(R0 * (1-p)))
  return(ifelse(val < 0, 0, ifelse(p >= P_crit(R0), 0, val)))
}

#r is the probability ratio of probability of getting vaccine related adverse outcome and disease related adverse outcome. r < 0 means percived risk of vaccination is lesser than that of disease itself
payoff <- function(P = 0.5, p = 0.5, r = 0.2, R0 = 2) {
  -(r*P) - (pi(p, R0)*(1-P))
}

#A sigma proportion is vaccinated with P probability and the rest with Q probability.
payoff_gain <- function(P, Q, sigma, rp = 0.2, rq = 0.2 ,R0 = 2) {
  
  #p is the overall proportion of vaccinated people
  p = sigma*P + ((1-sigma)*Q)
  
  EP = payoff(P = P, p = p, r = rp, R0 = R0)
  EQ = payoff(P = Q, p = p, r = rq,  R0 = R0)
  
  EP - EQ
}

#NE probability. If r > Pi(0), no one would vaccinate and csne_p will be 0.
csne_P <- function(r, R0 = 2) {
  val = 1 - (1/(R0*(1-r)))
  return(ifelse(val < 0, 0, ifelse(r >= pi(p = 0, R0), 0, val)))
}

P_crit <- function(R0 = 2) {
  p = 0
  if (R0 > 1){
    p = 1 - (1/R0)
  }
  p
}

csne_payoff <- function(r, R0) {
  P = csne_P(r = r, R0 = R0)
  p = P
  payoff(P,p,r,R0)
}
#As long as the percieved risk r = 0 and p < Pcritic, expected individual payoff ifcreases with P
```

If P is a Nash equilibrium, and everyone is currently playing P, then no one should change strategy. If P is convergently stable, then regardless of what strategy is most common in the population, individuals should start to play strategies closer to P, and ultimately adopt P.

```{r}
R0 <- 2
sigma <- 0.6
r <- 0.1
P <- seq(0,1, 0.01)
Q <- seq(0,1, by = 0.01)
ggplotly(expand_grid(P,Q) %>% mutate(gain = payoff_gain(P = P, Q = Q, sigma = sigma, rp = r, rq = r, R0 = R0)) %>% ggplot(aes(P, gain, col = Q)) + geom_line(aes(group = Q)) + geom_hline(yintercept = 0) + geom_vline(xintercept = csne_P(r, R0)))
```


```{r}
theme_set(theme_bw())
rs <- seq(0,1.5,by=0.01)
R0s <- c(2,5,10,20)
df <- expand.grid(rs,R0s) %>% rename(r = Var1, R0 = Var2)
df %>% mutate(p_star = csne_P(r = r, R0 = R0)) %>% ggplot(aes(r, p_star, group = R0)) + geom_line(aes(linetype = as.factor(R0)))
```

```{r}
rs <- seq(0,2,by=0.01)
R0s <- c(2,5,10, 20)
df <- expand.grid(rs,R0s) %>% rename(r = Var1, R0 = Var2)
df %>% mutate(p_fixed = csne_P(r = 1, R0 = R0), p_exp = csne_P(r = r, R0 = R0), delta = p_exp - p_fixed) %>% ggplot(aes(r, delta, lineltye = as.factor(R0))) + geom_line(aes(linetype = as.factor(R0)))
```


Analysis of vaccine scares: payoff gain, ΔE, and change in vaccine uptake, ΔP, after a shift in risk perception from r < π0 to r′ (see Table 1). For this figure, r = 0.1 and the proportion of individuals currently adopting the new CSNE is ε = 0 (corresponding to the start of a vaccine scare); the shapes of the curves are qualitatively similar for other values of r and ε.
```{r}
rs <- seq(0,1.5,by=0.01)
R0s <- c(2,5,10, 20)
sigma = 0
old_risk = .1

df <- expand.grid(rs,R0s) %>% rename(r = Var1, R0 = Var2)
df %>% 
    mutate(p_fixed = csne_P(r = old_risk, R0 = R0), 
           p_exp = csne_P(r = r, R0 = R0), 
           delta = p_exp - p_fixed, 
           e_gain = payoff_gain(P = p_exp, Q = p_fixed, sigma = sigma, rp = r, rq = r, R0 = R0)) %>% pivot_longer(., cols = c(delta, e_gain), names_to = 'nam', values_to = 'vals') %>% 
    ggplot(aes(r, vals, lineltye = as.factor(R0), col = nam)) + 
    geom_line(aes(linetype = as.factor(-R0)))
```


